---
title: "Analysis of the pupae data in rmarkdown"
author: "Remko Duursma"
date: "13 April 2016"
output: word_document
---
  

```{r echo=FALSE, eval=FALSE}
# INSTRUCTIONS - READ FIRST

# This file is a so-called 'rmarkdown' file. It combines text formatted in markdown,
# a common very easy-to-use language to format documents, with R code. To run this example,
# you need to have the following packages installed:
# knitr, rmarkdown, doBy

# This Rmd file can be converted to a Word document, where all R code is going to be executed for 
# us, and the results of the R code placed inside the document (including text, figures, etc).

# --> To do this, click the 'Knit Word' button that appears just above this window in Rstudio.
# If it fails, you probably do not have some of the packages installed that are mentioned above.
# If it works, a Word document will open up. Compare the output to the code below.
# --> Also try to convert this document into an HTML document, which displays very nicely in a browser.
# To do this, select 'Knit HTML' from the menu that, by default, displays 'Knit Word' (click on the small triangle)


# The R code can appear in so-called 'code chunks', or even as code within text.
# This text also appears in a code chunk, and note that I have set two options:
# echo=FALSE (this will not appear in the Word document we are constructing), and
# eval=FALSE (this code will not even be executed, handy when a code chunk contains code that
# does not work or takes too long to run!)

# For more options and instructions, see rmarkdown.rstudio.com (for all things markdown, which affects
# the formatting of text outside code chunks), and http://yihui.name/knitr/options/ (for all options that you can give to the code chunks).
```






# Introduction

This dataset is from an experiment where larvae were left to feed on *Eucalyptus* leaves,
in a glasshouse that was controlled at two different levels of temperature and CO~2~ concentration.
After the larvae pupated (that is, turned into pupae), the body weight was measured, as well as the cumulative 'frass' (larvae excrement) over the entire time it took to pupate.

Note the formatting above, we can have *italic*, or even **bold** text, super^script^ and sub~script~.


  
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# This is an example of a code block which you do not want to include in
# the final document (echo=FALSE)
# We also set message=FALSE, and warning=FALSE, to avoid inclusion of messages
# and warnings in the final document.

pupae <- read.csv("pupae.csv")
pupae$CO2_treatment <- as.factor(pupae$CO2_treatment)

library(doBy)
puptab <- summaryBy(Frass + PupalWeight ~ CO2_treatment + T_treatment,
                    data=pupae, FUN=c(mean, sd), na.rm=TRUE)

# It is more convenient to reorder the dataframe to have sd and mean
# together.
puptab <- puptab[,c("CO2_treatment","T_treatment","Frass.mean","Frass.sd","PupalWeight.mean","PupalWeight.sd")]

# Rename
names(puptab) <- c("CO2","T","Frass","SD","PupalWeight","SD")

# Round if the variable is 'double' (i.e. not integer or character)
puptab <- as.data.frame(lapply(puptab, function(x)if(is.double(x))round(x,2)else x))
```

# Results

In rmarkdown, results from statistical analyses are shown in grey boxes. For example,

```{r}
lmfit <- lm(Frass ~ PupalWeight * Gender, data=pupae)
anova(lmfit)
```

The above result simply shows the text output of the anova, which is useful in itself but not very pretty. To make a quick pretty table, we can use the `kable` function from the knitr package.

```{r}
library(knitr)
kable(anova(lmfit))
```

The above table is converted to an actual Word table, so you can change formatting in the usual way.

## Inline results

Another very useful feature of rmarkdown is that you can include results within text. For example, I calculated the standard deviation of Frass to be `r sd(pupae$PupalWeight)`, and the R^2^ of the regression shown above was `r summary(lmfit)$r.squared`. 


# Figures

It is straightforward to include code and output in markdown, this includes figures. Unfortunately we cannot add automatic captions to the figures.

```{r dpi=300}
# Example where we show the code to make the figure.
# Note the code chunk option 'dpi=300', this increases the resolution of the figure,
# which is otherwise quite low.
with(pupae, plot(Frass, PupalWeight, pch=c(1,19)[CO2_treatment],
                 col=c("blue","red")[T_treatment]))
```


```{r echo=FALSE, dpi=300}
# Example where we do not show the code
boxplot(Frass ~ Gender, data=pupae, xlab="Gender", ylab="Frass")
```

The code for the above figure is not shown in the Word document.

# Tables

Including tables with results into Word has always been a struggle. Not anymore. Below are two options to include tables in Word based on a summary table made with `summaryBy`.

The following table is made with the kable function from the knitr package. Note how we include a caption for the table as an argument to kable.

```{r echo=TRUE, results="asis"}
library(knitr)

kable(puptab, caption="Table 1. Summary stats for the pupae data.")
```

The last version uses the pander function from the package with the same name.

```{r results="asis", echo=TRUE, eval=FALSE}
# This code chunk does not execute by default, install pander and set eval=TRUE first
library(pander)
pander(puptab)
```

